The Causes of Damage in Toronto Fire Incidents

### 2024-09-27 paper/paper.qmd
Step 4.2: finished Checks(E.1.2) according to book

---
title: "The Causes of Damage in Toronto Fire Incidents"
author: 
  - David Qi
thanks: "Code and data are available at: https://github.com/UTDQi/Fire_Incidents"
date: today
date-format: long
abstract: "This paper discusses some basic findings around the monetary and lives loss in Toronto fire incidents. By using data from open data Toronto and choosing the relevant entries, we have identified possible factors that influences monetary and lives loss in fire incidents. We identified unattended ignition sources as a significant risk factor for civilian casualties, highlighting the need for public awareness campaigns to educate communities about these potential hazards and noted that fires with undetermined causes resulted in greater property loss. Interestingly, we observed that increasing the number of responding personnel beyond 90 led to a reduction in civilian casualties, which contrasts with our initial expectations."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

Fire incidents are a major threat to urban lifestyles, causing substantial loss of life and property. In recent years, there has been an increase in number of fire incidents reported in Toronto (@fig-years), which highlights the importance of identifying trends and patterns in historical data. These trends can provide helpful insights on causation of casualties and monetary loss, which may further enable us to find areas of improvement for our fire services. As a simple comparison, we also used the fire incident data in Detroit from city of Detroit open data portal(@Detroit).

@@

The remainder of this paper is structured as follows: We first give a brief overview of the dataset under [Data](#sec-data) section.Providing some basic information. Then, we use graphs and models to present the results we find from our data in the [results](#results) section. We then discusses the possible interpretations of our find further in the [discussion](#discussion) section. While also pointing our the weaknesses and areas for further investigations in the [Weaknesses and next steps](#next-step) section.

@@

From @fig-cause-compare-dollar and @fig-cause-compare-casualties,we see that among the top five causes of fires we identified,  each of them are substantially different in the damage they create.
Unattended ignition source is the most dangerous in causing death, but on average cause relatively less damage to properties. This suggests that causalities and property damage are different in nature and have to be treated differently. This observation also captures the nature of unattended unattended ignition sources, which often catch individuals off-guard, leading to a higher risk. 
The vast difference also highlights the importance of identifying the cause of fire, which may help us lower potential  property damage. And help us in understanding how to treat the fire.
The mean dollar loss for all causes of fire is 38,105.88, while the mean dollar loss for undetermined causes is significantly higher at 63,034. This indicates that fires of undetermined origin often result in more severe damage, possibly due to delayed detection and complex fire behavior. Undetermined fires are more likely to involve unusual circumstances, complicating both firefighting efforts and post-incident analysis. Understanding these patterns could be crucial for improving firefighting strategies.

@@

  The number of responding personnel, such as firefighters and emergency medical teams, plays a critical role in determining the outcomes of fire incidents involving casualties and property damage. From our analysis, we observed an unexpected decline in the number of casualties when the number of responding personnel exceeds 90, highlighting a key area for further study. Since the fire incident site involves complex decision-making and coordination, it is crucial to avoid both overstaffing and understaffing in order to maintain an effective and manageable response team.

@@

  We have noted that examining the existing organization of incidents where the number of responding personnel exceeds 90 may reveal insights into the reasons behind the observed decrease in casualties. Understanding the factors that contribute to this reduction can contribute to better resource allocation and emergency response management.

>>>>

---
title: "The Causes of Damage in Toronto Fire Incidents"
author: 
  - David Qi
thanks: "Code and data are available at: https://github.com/UTDQi/Fire_Incidents"
date: today
date-format: long
abstract: "This paper discusses some basic findings around the monetary and lives loss in Toronto fire incidents. By using data from open data Toronto and choosing the relevant entries, we have identified possible factors that influences monetary and lives loss in fire incidents. We identified unattended ignition sources as a significant risk factor for civilian casualties, highlighting the need for public awareness campaigns to educate communities about these potential hazards. Interestingly, we observed that increasing the number of responding personnel beyond 90 led to a reduction in civilian casualties, which contrasts with our initial expectations."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

Fire incidents are a major threat to urban lifestyles, causing substantial loss of life and property. In recent years, there has been an increase in number of fire incidents reported in Toronto (@fig-years), which highlights the importance of identifying trends and patterns in historical data. These trends are toward conclusions on causation of casualties and monetary loss, which may further enable us to find areas of improvement for our fire services. As a simple comparison, we also used the fire incident data in Detroit from city of Detroit open data portal(@Detroit).By studying these data, we find that fire with undetermined cause on average create more property damage, unattended ignition sources is the greatest threat to civilians among the frequent fire causes, and personnel responding can lower the civilian causality when exceeding 90.

@@

The remainder of this paper is structured as follows: We first give a brief overview of the dataset under [Data](#sec-data) section.Providing some basic information. Then, we use graphs and models to present the results we find from our data in the [Results](#results) section. We then discusses the possible interpretations of our find further in the [Discussion](#discussion) section. While also pointing our the weaknesses and areas for further investigations in the [Weaknesses and next steps](#next-step) section.

@@

From @fig-cause-compare-dollar and @fig-cause-compare-casualties,we see that among the top five causes of fires we identified,  each of them are substantially different in the damage they create.
Unattended ignition source is the most dangerous in causing death, but on average cause relatively less damage to properties. This suggests that causalities and property damage are different in nature and have to be treated differently. This observation also captures the nature of unattended unattended ignition sources, which often catch individuals off-guard, leading to a higher risk. 
The vast difference also highlights the importance of identifying the cause of fire, which may help us lower potential  property damage. And help us in understanding how to treat the fire.
The mean dollar loss for all causes of fire is 38,105.88, while the mean dollar loss for undetermined causes is significantly higher at 63,034. This indicates that fires of undetermined origin often result in more severe damage, possibly due to delayed detection and complex fire behavior. Undetermined fires are more likely to involve unusual circumstances, complicating both firefighting efforts and post-incident analysis. Understanding these patterns could be key for improving firefighting strategies.

@@

  The number of responding personnel, such as firefighters and emergency medical teams, plays a critical role in determining the outcomes of fire incidents involving casualties and property damage. From our analysis, we observed an unexpected decline in the number of casualties when the number of responding personnel exceeds 90, highlighting a key area for further study. Since the fire incident site involves complex decision-making and coordination, it is vital to avoid both overstaffing and understaffing in order to maintain an effective and manageable response team.

@@

  We have noted that examining the existing organization of incidents where the number of responding personnel exceeds 90 may bring understandings into the reasons behind the observed decrease in casualties. Understanding the factors that contribute to this reduction can contribute to better resource allocation and emergency response management.
